On $K_2$-analogue of Serre problem for $\SL_4$

### 2018-06-05 paper.tex
Introduced set of adjacency classes

@@

@@

\subsection{$\sigma$-homomorphisms}
%Our notation and conventions follows~\cite[\S~4]{Vav09}.
Let $\Phi$ be an irreducible root system with some fixed basis of simple roots $\Pi = \{\alpha_1, \ldots, \alpha_\ell\}$.
For a root $\alpha\in\Phi$ we denote by $m_i(\alpha)$ the $i$-th coefficient in the expansion of $\alpha$ in $\Pi$,
 i.\,e. $\alpha = \sum_{i=1}^n m_i(\alpha) \alpha_i$.

We denote by $\Phi^\vee$ the \emph{dual root system of $\Phi$} consisting of vectors $\alpha^\vee = 2\alpha/(\alpha, \alpha)$, $\alpha\in \Phi$.
As usual, $P(\Phi^\vee)$ denotes the lattice spanned by the \emph{fundamental weights $\varpi_i$}.
Recall that $\varpi_i$ are uniquely determined by relations $\langle\varpi_i, \alpha_j^\vee \rangle = (\varpi_i, \alpha_j) = \delta_{ij}.$

Notice that for $\varpi \in P(\Phi^\vee)$ and $\beta \in \ZZ \Phi$ one has $(\varpi, \beta) \in \ZZ $.
Consequently, for $\varepsilon \in R^*$ and $\varpi \in P(\Phi^\vee)$ the formula $\chi_{\varpi, \varepsilon}(\beta) = \varepsilon ^ {(\varpi, \beta)}$
defines a character $\chi_{\varpi, \varepsilon} \in \Hom(\ZZ \Phi, R^*)$.

@@

In the sequel we assume that $i$ is a vertex of the Dynkin diagram $D(\Phi)$ of $\Phi$ with the property that $m_i(\alpha_{\max})=1$.
Denote by $\mathcal{X}_i$ (resp. $\mathcal{X}^i$) the subsets of $\mathcal{X}_{\Phi, A[t]}$ consisting of those generators $x_{\alpha}(\xi)$ of $\St(\Phi, A[t])$ for which 
$m_i(\alpha) > 0$ (resp. $m_i(\alpha)<0$) implies that $\xi \in A[t]$ is divisible by $t$.
% either $i\neq k$ (resp. $j\neq k$) or $\xi\in R[t]$ is divisible by $t$.
Denote by $N_i$ and $N^i$ the subgroups of $\St(\Phi, A[t])$ generated by $\mathcal{X}_i$ and $\mathcal{X}^i$, respectively.
%Our goal is to construct mutually inverse morphisms $\sigma_k$, $\sigma_k^{-1}$ fitting into the following diagram.

\begin{dfn} A pair of mutually inverse maps $(\sigma_i^+, \sigma_i^-)$ between the groups $N^i$ and $N_i$ will be called a {\it $\sigma$-pair corresponding to the vertice $i\in D(\Phi)$},
 if these maps fit into the following diagram.
\begin{equation} \label{eq:sigma-diagram}
\xymatrix{ N_k\ar@{-->}[r]^{\sigma_i^+}\ar[d] & N^k \ar[d] \ar@{-->}[r]^{\sigma_i^-} & N_k \ar[d] \\ 
          \St(\Phi, R) \ar@<-0.0ex>[r]_{\beta_i^+} & \St(\Phi, R) \ar@<-0.0ex>[r]_{\beta_i^-} & \St(\Phi, R)} 
\end{equation}
In the above diagram the vertical arrows are canonical embeddings of $N_k$ and $N^k$ into $\St(\Phi, A[t])$ postcomposed with 
 the map of Steinberg groups induced by the embedding $A[t] \hookrightarrow R$.
%As before, we denote by $R$ the ring $A[t, t^{-1}]$ of Laurent polynomials over $A$.
         % \GL(n, R) \ar@<-0.0ex>[r]_{(-)^{\delta_k}} & \GL(n, R) \ar@<-0.0ex>[r]_{{}^{\delta_k}(-)} & \GL(n, R)
\end{dfn}

@@

\subsection{Construction of \texorpdfstring{$\sigma$}{sigma}-morphisms for $SL_4$}
Let $R$ be a local ring and assume that $n\geq 4$.
For $1\leq k\leq n$ denote by $\delta_k$ the $n\times n$ matrix of $\GL(n, R[t, t^{-1}])$ which differs from the unit matrix 
 only by the element $t$ on the $k$-th place of its diagonal.
Recall from~\cite[Corollary~4]{Ka77} that there exists unique automorphisms $\beta_{\delta_k}$, $\beta_{\delta_k^{-1}}$ of $\St(n, R[t, t^{-1}])$
 modeling conjugation with matrices $\delta_k$ and $\delta_k^{-1}$ (see diagram~\eqref{eq:sigma-diagram} below)

>>>>

\newcommand{\rA}{\mathsf{A}}
\newcommand{\rB}{\mathsf{B}}
\newcommand{\rC}{\mathsf{C}}
\newcommand{\rD}{\mathsf{D}}
\newcommand{\rE}{\mathsf{E}}
\newcommand{\rF}{\mathsf{F}}
\newcommand{\rG}{\mathsf{G}}

@@

Our notation and conventions follows~\cite[\S~4]{Vav09}.
Let $\Phi$ be an irreducible root system with some fixed basis of simple roots $\Pi = \{\alpha_1, \ldots, \alpha_\ell\}$.
For a root $\alpha\in\Phi$ we denote by $m_i(\alpha)$ the $i$-th coefficient in the expansion of $\alpha$ in $\Pi$,
 i.\,e. $\alpha = \sum_{i=1}^n m_i(\alpha) \alpha_i$.

We denote by $\Phi^\vee$ the \emph{dual root system of $\Phi$} consisting of coroots $\alpha^\vee = 2\alpha/(\alpha, \alpha)$, $\alpha\in \Phi$.
As usual, $P(\Phi^\vee)$ denotes the lattice spanned by the \emph{fundamental weights $\varpi_i$},
 that are uniquely determined by relations $\langle\varpi_i, \alpha_j^\vee \rangle = (\varpi_i, \alpha_j) = \delta_{ij}.$ %TODO: Mistake?

@@

\subsection{$\sigma$-pairs}
Notice that for $\varpi \in P(\Phi^\vee)$ and $\beta \in \ZZ \Phi$ one has $(\varpi, \beta) \in \ZZ $.
Consequently, for $\varepsilon \in R^*$ and $\varpi \in P(\Phi^\vee)$ the formula $\chi_{\varpi, \varepsilon}(\beta) = \varepsilon ^ {(\varpi, \beta)}$
defines a character $\chi_{\varpi, \varepsilon} \in \Hom(\ZZ \Phi, R^*)$.

@@

In the sequel we assume that $i$ is a vertex of the Dynkin diagram $D(\Phi)$ of $\Phi$ with the property that $m_i(\alpha_{\max})=1$ (of course, this implies that $\Phi$ is not of type $\rE_8, \rF_4$ or $\rG_2$).
Denote by $\mathcal{X}_i$ (resp. $\mathcal{X}^i$) the subsets of $\mathcal{X}_{\Phi, A[t]}$ consisting of those generators $x_{\alpha}(\xi)$ of $\St(\Phi, A[t])$ for which 
$m_i(\alpha) > 0$ (resp. $m_i(\alpha)<0$) implies that $\xi \in A[t]$ is divisible by $t$.
% either $i\neq k$ (resp. $j\neq k$) or $\xi\in R[t]$ is divisible by $t$.
Denote by $N_i$ and $N^i$ the subgroups of $\St(\Phi, A[t])$ generated by $\mathcal{X}_i$ and $\mathcal{X}^i$, respectively.
%Our goal is to construct mutually inverse morphisms $\sigma_k$, $\sigma_k^{-1}$ fitting into the following diagram.

\begin{dfn} A pair of mutually inverse maps $(\sigma_i^+, \sigma_i^-)$ between the groups $N^i$ and $N_i$ will be called a {\it $\sigma$-pair corresponding to the vertex $i\in D(\Phi)$},
 if these maps fit into the following diagram.
\begin{equation} \label{eq:sigma-diagram}
\xymatrix{ N_i\ar@{-->}[r]^{\sigma_i^+}\ar[d] & N^i \ar[d] \ar@{-->}[r]^{\sigma_i^-} & N_i \ar[d] \\ 
          \St(\Phi, R) \ar@<-0.0ex>[r]_{\beta_i^+} & \St(\Phi, R) \ar@<-0.0ex>[r]_{\beta_i^-} & \St(\Phi, R)} 
\end{equation}
In the above diagram the vertical arrows are canonical embeddings of $N_k$ and $N^k$ into $\St(\Phi, A[t])$ postcomposed with 
 the map of Steinberg groups induced by the embedding $A[t] \hookrightarrow R$.
%As before, we denote by $R$ the ring $A[t, t^{-1}]$ of Laurent polynomials over $A$.
         % \GL(n, R) \ar@<-0.0ex>[r]_{(-)^{\delta_k}} & \GL(n, R) \ar@<-0.0ex>[r]_{{}^{\delta_k}(-)} & \GL(n, R)
\end{dfn}

For the remainder of this subsection our main goal will be to verify that an analogue of the group decomposition from \cite[Lemma~3.1f]{Tu83} holds for $\St(\Phi, A)$
 provided $\St(\Phi, R)$ admits at least one $\sigma$-pair.

Denote by $\widetilde{W}$ the subgroup of $\St(\Phi, A)$ generated by elements $w_\alpha(1)$.
The subgroup $\widetilde{W}$ acts on $\St(\Phi, A)$ via conjugation, for example, 
 the action on the generators of $\mathcal{X}_{\Phi, A}$ can be expressed via the formula ${}^{w_\alpha(1)} \!x_{\beta}(\xi) = x_{\sigma_{\alpha}(\beta)} ( \pm \xi)$.
It is easy to see that the action of $\widetilde{W}$ on the set of root subgroups $\{X_\alpha \mid \alpha \in \Phi \}$ of $\St(\Phi, R)$ coincides with the obvious
 action of the Weyl group $W(\Phi)$ on it. Since $N_i$ and $N^i$ are generated by root subgroups we immediately obtain the following statement.
\begin{lemma} The set of elements in the orbit of $N_i$ (resp. $N^i$) under the conjugation action of $\widetilde{W}$ 
is in bijective correspondence with 
the set of adjacency classes of $W(\Phi)$ with respect to its Levi subgroup $W_i = W(\Delta_i)$. %TODO: Unsure whether it is true at all :)
\end{lemma}
In the sequel we denote by $\Sigma$ the set of elements in $\widetilde{W}$ whose projections in $W(\Phi)$ form the set of representatives for
 right adjacency classes of $W(\Phi)$ with respect to $W_i$. %We can assume these representatives to be short

@@

\subsection{Construction of \texorpdfstring{$\sigma$}{sigma}-morphisms for $\SL_4$}
Let $R$ be a local ring and assume that $n\geq 4$.
For $1\leq k\leq n$ denote by $\delta_k$ the $n\times n$ matrix of $\GL(n, R[t, t^{-1}])$ which differs from the unit matrix 
 only by the element $t$ on the $k$-th place of its diagonal.
Recall from~\cite[Corollary~4]{Ka77} that there exists unique automorphisms $\beta_{\delta_k}$, $\beta_{\delta_k^{-1}}$ of $\St(n, R[t, t^{-1}])$
 modeling conjugation with matrices $\delta_k$ and $\delta_k^{-1}$ (see diagram~\eqref{eq:sigma-diagram} below)
